ODE and Diffusion Models of Tumor Growth

### 2024-03-13 paper.tex
3 pages done

\end{document}

>>>>

This equation states that the rate of change of the concentration of oxygen plus the rate of usage of the oxygen is equal to the diffusion equation.
Note that from the assumptions above, the concentration is held constant, so the equation can be further simplified to:

\begin{equation}
a = k \* \frac{d^2y}{dx^2}
\end{equation}

The solution of the equation is

\begin{equation}
y = \frac{ax^2}{2k} + bx + y_0
\end{equation}

where $b$ is a constant.
This can be verified by taking the derivative of this equation twice:

$ \frac{dy}{dx} = \frac{a}{k}x + b$,

\vspace*{0.25cm}

$ \frac{d^2y}{dx^2} = \frac{a}{k}$
\vspace*{0.25cm}

To find the value of b that satisfies these equations, we assume that there must be some point 
very far away from the source where the concentration is zero and the diffusion of oxygen is not occuring,
so $\frac{dy}{dx}$ is also equal to zero.

We define $x'$ as the minimum distance where the concentration is equal to zero and then set these two equations equal to zero and solve:

\vspace*{0.25cm}
$ 0 = \frac{a{x'}^2}{2k} + bx' + y_0 $
\vspace*{0.25cm}

$ 0 = \frac{ax'}{k} + b \Leftrightarrow b = -\frac{ax'}{k}$
\vspace*{0.5cm}

Substituting back into the first equation:
\vspace*{0.5cm}

$ 0 = \frac{a{x'}^2}{2k} - \frac{a{x'}^2}{k} + y_0 \Leftrightarrow y_0 = \frac{a{x'}^2}{2k}$

So 
\begin{equation}
    x' = \sqrt{\frac{2ky_0}{a}}
\end{equation}
(the maximum permeation distance of the oxygen from the source)

and

\begin{equation}
    b = \sqrt{\frac{-2ay_0}{k}}
\end{equation}

We can now solve to find how much oxygen has been absorbed by the tissue: $ \int_{0}^{x'} y dx $, where y is found in
equation (3).

We get the integral:

\begin{equation}
    Y = \int_{0}^{\sqrt{\frac{2ky_0}{a}}} \frac{ax^2}{2k} + bx + y_0 dx
\end{equation}

To solve (7), we integrate and then simplify:

\vspace*{0.25cm}

$Y = \frac{ax^3}{6k} - \frac{x^2}{2}\sqrt{\frac{2ay_0}{k}} + y_0x \Bigg|_0^{\sqrt{\frac{2ky_0}{a}}}$
\vspace*{0.25cm}

$ = \sqrt{2ky_0} \frac{y_0}{3\sqrt{a}} - \frac{ky_0\sqrt{\frac{2ay_0}{k}}}{a} + y_0\sqrt{\frac{2ky_0}{a}}$
\vspace*{0.25cm}

\begin{equation}
    = y_0 \frac{x'}{3}
\end{equation}

This result (which Hill also derived) shows that the maximum absorption of oxygen in a tissue that is $x'$ thick
is one-third the full amount of oxygen that is diffused. 
\end{document}
